---
title: "Practical Machine Learning - Prediction Assignment"
author: "Volety Sai Sirisha"
date: "11 October 2020"
output: html_document
---

## Overview
The aim of this project is to predict the manner in which participants perform a exercises. The data comes from http://groupware.les.inf.puc-rio.br/har wherein 6 participants were asked to perform the same set of exercises correctly and incorrectly.

## Training Data and Testing Data
The training data is taken from:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The testing data is taken from:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

## Load Packages
```{r package}
set.seed(1000)
library(caret)
library(ggplot2)
library(corrplot)
library(rpart)
library(randomForest)
library(rattle)
library(rpart.plot)
```

## Load and Partition Data
```{r load}
#Loading the training data
train_data <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"))
#Loading the test data
test_data <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"))

label <- createDataPartition(train_data$classe, p = 0.7, list = FALSE) #partition data
training <- train_data[label, ]
testing <- train_data[-label, ]
```

## Cleaning data
From the 160 variables there are some NA values and Zero Variance that have to be removed.
```{r clean data}
nzv <- nearZeroVar(training) #Zero variance 
training <- training[ ,-nzv]
testing <- testing[ ,-nzv] 

label <- apply(training, 2, function(x) mean(is.na(x))) > 0.95  #NA values
training <- training[, -which(label, label == FALSE)]
testing <- testing[, -which(label, label == FALSE)]

training <- training[ , -(1:5)]  #Remove the other 5 variables
testing <- testing[,-c(1:5)]

dim(training) #Check the dimension of the train data
dim(testing)  #Check the dimension of the test data 
```

## Exploratory Analysis
From the graph we observe that the darker gradient corresponds to having higher correlation.
```{r analysis}
corrMat <- cor(training[,-54])
corrplot(corrMat, method = "color", type = "upper", tl.cex = 0.8, tl.col = rgb(1,0,1))
```

## Training the Models using different methods
We will be using 2 different model to train the data to observe the affects. The 2 types are : Decision Tree, Random Forests.
A) Decision Tree
```{r Decision Tree}
model_DecT <- rpart(classe ~ ., data = training, method = "class")
predict_DecT <- predict(model_DecT, testing, type = "class")
confMat_DecT <- confusionMatrix(predict_DecT, testing$classe)
confMat_DecT
```
B)Random Forest
```{r random forest}
ctrl <- trainControl(method = "cv", number = 3, verboseIter=FALSE)
model_RanF <- train(classe ~ ., data = training, method = "rf", trControl = ctrl)
predict_RanF <- predict(model_RanF, testing)
confMat_RanF <- confusionMatrix(predict_RanF, testing$classe)
confMat_RanF
```

We see that the accuracy of Random Forest is better than Decision Tree. Therefore we use Random forest to predicts the results.

## Predicting the test output
```{r output}
predict_RanF <- predict(model_RanF, test_data)
predict_RanF
```


